\section{Solving the 1D diffusion equation with a delta-sink and absorbing boundaries}

\subsection{Motivation}

In this paper a 1D greensfuction with a delta-sink and two absorbing boundaries is derived. Furtheremore, other quantaties important for $\GFRD$, such as the survival probability, the c.d.f. in space and the fluxes at the boundarie and the sink are given. Finally we will consider the convergence of the series and other properties important for the numeric implementation of the functions.

\subsection{Introduction}

To simulate the 1D diffussion of a protein allong the DNA in the pressence of a promotor site with $\GFRD$, we need a new type of domain. This domain contains a delta-sink which simulates the promotor site. In contrast to the original domains which have an absorbing boundary condition, particles are not reflected in the case a particle doesn't react, but instead just diffuse over the reactive spot. This reactive spot will generate a probability current leaving the domain proportianal to the rate constant $k$ and the probability of a particle being at that specific site. This current represents the particles wich bind to the promotor. Other sink terms are the two absorbing boundaries at both sites of the domain. The fluxes at these three point are used in $\GFRD$ to determine the event type at the next event time.

\subsection{Obtaining the Green's function}

\subsubsection{Conditions}

We want to find the Green's function $p(x,t|x_0,\tau)$ for a particle with diffussion constand $D$, obbaying the differential equation,
\begin{equation}
 \frac{\partial p(x,t|x_0,\tau)}{\partial t} = D \, \nabla ^2 p(x,t|x_0,\tau) - k \, \delta (x - x_s) p(x,t|x_0,\tau)
 \label{DE_I}
\end{equation}
on a one dimensional domain with absorbing boundary conditions at $L_l$ (left boundary) and $L_r$ (right boundary),
\begin{eqnarray}
 p(L_l,t|x_0,\tau) & = & 0 \\
 p(L_r,t|x_0,\tau) & = & 0
\label{BC_I}
\end{eqnarray}
and a unit instantaneous source at $x_0$ as initial condition
\begin{equation}
 p(x,t|x_0,\tau) = \delta(x-x_0) \delta(t-\tau).
\label{IC_I}
\end{equation}

\subsubsection{Solving in Laplace space}

Solving equation (\ref{DE_I}) directly is difficult due to the term containing both the delta function and the function we want to solve for. We can however split up the domain $[L_l,L_r]$ in to the sub-domains $[L_l,x_s]$ and $[x_s,L_r]$, with the delta sink in in between them. The delta-sink is now introduced into the problem by the continuity conditions which have to hold between the two domains. 

\subsubsection{The continuity conditions}

The continuity equation for our problem is

\begin{equation}
 \nabla \vec{J}(x,t|x_0) + \partial_t p(x,t|x_0) = -k \delta( x - x_s ) p(x,t|x_0),
\end{equation}
which means that probability density is conserved everywhere except at the postition of the sink $x_s$, where is 'dissapears' with a rate $k \delta ( x - x_s ) p$. We can simplify this condition by first plugging in Fick's law for the probability current $J$, and integrating over a small region $[a - \epsilon,a + \epsilon]$ around the sink, 
\begin{gather}
\int_{x_s - \epsilon}^{x_s + \epsilon} \partial_x \left( -D \partial_x p(x,t|x_0)  \right) \, dx + \partial_t \int_{x_s - \epsilon}^{x_s + \epsilon} p(x,t|x_0) \, dx = \nonumber \\
-k \int_{a - \epsilon}^{a + \epsilon} \delta( x - x_s ) p(x,t|x_0) \, dx \nonumber \\
\left[ -D \partial_x p(x,t|x_0)  \right]_{x_s - \epsilon}^{x_s + \epsilon} + \partial_t \left. P(x,t|x_0) \right|_{x_s - \epsilon}^{x_s + \epsilon} = -k p(x_s,t|x_0) \nonumber \\
J_r(x_s + \epsilon,t|x_0) - J_l(x_s - \epsilon,t|x_0) + \partial_t \frac{P(x_s + \epsilon,t|x_0) - P(x_s - \epsilon,t|x_0)}{2 \epsilon} 2\epsilon = \nonumber \\ 
-k p(x_s,t|x_0). \nonumber \\
\end{gather}
were we multiplied the second term on the left hand side with $2\epsilon/2\epsilon$. Now taking the limit $\epsilon \rightarrow 0$,
\begin{gather}
 \lim_{\epsilon \to 0} J_r(x_s + \epsilon,t|x_0) - J_l(x_s - \epsilon,t|x_0) + \partial_t p(x_s,t|x_0) 2\epsilon = -k p(x_s,t|x_0) \nonumber = \\
 J_r(x_s,t|x_0) - J_l(x_s,t|x_0) = -k p(x_s,t|x_0).
\end{gather}
Clearly, the sink causes a discontinuity in the flux from the left to the right domain $J_l$ and vice versa flux $J_r$, at the boundary. We use this result for the continuity conditions between the two domains, which are:

(I) The solution for the left domain $p_l$ has to equal the solution for the right domain $p_r$ at $x_s$,
\begin{equation} 
 p_l(x_s,t|x_0,\tau) = p_r(x_s,t|x_0,\tau) 
\end{equation}.
(II) The flux from the left domain at $x_s$ should equal the flux from the right domain minus the flux which left via the sink,
\begin{equation} 
 \left. \nabla p_l(x,t|x_0,\tau) \right|_{x_s} = \left. \nabla p_r(x,t|x_0,\tau) \right|_{x_s} - \frac{k}{D} \, p_{l}(x_s,t|x_0,\tau).
\end{equation}

\subsubsection{Ansats equations}

The difficult problem of solving (\ref{DE_I}) has now been simplified to solving the well known heat equation in both domains. To simplify even more we let the Laplace operator act on the time domain of the heat equation. The time variable $t$ is replaced by the Laplace variable $s$ in all functions (Laplace transformed functions are denoted by a bar on top $\bar{f}$) and the time derivative in the heat equation changes in a multiplicative factor;
\begin{equation}
 s \, \bar{p}(x,s|x_0) = D \, \nabla^2 \bar{p}(x,s|x_0).
\end{equation}
Furthermore we assume the delta-sink to be positioned at the origin $x_s=0$. Together with the boundary conditions (\ref{BC_I}) and the initial condition (\ref{IC_I}), were we assume the position of the initial delta peak to be in the right domain, we can write down an Ansatz in Laplace space for the two domains.
For the right domain,
\begin{equation}
 \bar{p}_r(x,x_0,q) = \frac{1}{2 D q} e^{-q |x-x_0|} + A \, \mathrm{cosh} (q x) + B \, \mathrm{sinh} (q x)
 \label{Anz_R}
\end{equation}
and the left domain
\begin{equation}
 \bar{p}_l(x,x_0,q) = E \, \mathrm{cosh} (q x) + F \, \mathrm{sinh} (q x),
\label{Anz_L}
\end{equation}
where the unknowns $A,B$ and $E,F$ can be functions of $q$, and $q=\sqrt{s/D}$. By filling in the Ansatz into the boundary and continuity conditions we find the four unknowns. Plugging these back into the Ansatz and simplifying, the solution in Laplace space becomes
\begin{equation}
 \bar{p}_r(x,q|x_0) = -2 \, \frac{\mathrm{sinh} \, q(L_r - x)}{2 D q} \left( \frac{Dq \, \mathrm{sinh} \, q(L_l + x_0) + k \, \mathrm{sinh} \, q L_l \,\, \mathrm{sinh} \, q x_0}{Dq \, \mathrm{sinh} \, q(L_r + L_l) + k \, \mathrm{sinh} \,\, q L_l \, \mathrm{sinh} \, q L_r} \right)
 \label{SOL_R}
\end{equation}

\begin{equation}
 \bar{p}_l(x,q,|x_0) = -\frac{\mathrm{sinh} \, q(L_l + x) \,\, \mathrm{sinh} \, q(L_r - x_0)}{Dq \, \mathrm{sinh} \, q(L_r + L_l) + k \, \mathrm{sinh} \, q L_l \,\, \mathrm{sinh} \, q L_r}
\label{SOL_L}
\end{equation}

\subsubsection{Transforming the Laplace solution to the time domain}

In order to transform our solution in Laplace space back to the time domain solving the following integral is paramount,
\begin{equation}
 \frac{1}{2 \pi i} \int_{-\infty i + \gamma}^{+\infty i + \gamma} \, e^{s t} \, \bar{p}(x,s|x_0) \, ds.
 \label{INT_I}
\end{equation}
Our solution in Laplace space is not of a simple form which can be looked up in a transform table, neither can we use the primitive of the integrand and fill in the integration boundaries; finding the primitive is beyond our capabilities. Therefore we have to reside to residual integration. 

When we integrate a function $f$ of a complex variable $z$ counterclockwise around a closed contour $\gamma$, this is equal to the sum over all residuals at positions $z = c_i$ of $f$ within $\gamma$
\begin{equation}
 2 \pi i\,\sum_{i=1}^{N}\,\mathrm{Res}(f,c_i) = \oint_\gamma f(z)\,dz.
 \label{RES_I}
\end{equation}

Investigating our solutions (\ref{SOL_R}) and (\ref{SOL_L}), we find that they are rational functions where both the numerator and denominator are holomorphic. We denote the numerator by $g_{r/l}(x,q|x_0)$ and the denominator by $h(x,q|x_0)$, where the subscript $r/l$ of the numerator indicates the solution for the right or the left domain. Both solutions have the same denominator apart from the factor $2 D q$ in the right domain, which we ignore for now because it cancels later on. Our solutions $p_{r/l}$ can hence be written as
\begin{equation} 
 \bar{p}_{r/l}(x,q|x_0) = \frac{g_{r/l}(x,q|x_0)}{h(q)}.
\end{equation}
From the theory of complex integration we know that the residues of these functions can be found via,
\begin{equation} 
 \mathrm{Res} \left( \frac{g_{r/l}(x,q|x_0)}{h(q)},\beta_i \right) = \left. \frac{g_{r/l}(x,q|x_0)}{\partial_q h(q)} \right|_{q = \beta_i}
 \label{RES_II}
\end{equation}
where the $\beta_i$ 's are the positions of the simple poles and are the roots of the denominator $h$,
\begin{equation} 
 D\beta_i \, \mathrm{sinh} \, \beta_i (L_r + L_l) + k \, \mathrm{sinh} \, \beta_i L_l \, \mathrm{sinh} \, \beta_i L_r = 0.
 \label{ROOT_I}
\end{equation}

The above equation only holds for real $q$ when $q = 0$, but this is not a simple pole of $\bar{p}$ because $g(x,q|x_0) = 0$ at this point. The relation does have however, an infinite amount of solutions when $q$ is purely imaginary. Remember that $q = \sqrt{s/D}$, which means that there are only poles for real $s < 0$. Looking at equation (\ref{INT_I}), the negative $s$ will give us a solution in the time domain which is monotonically decaying. This is obvious because for $t > 0$ we only have sinks and no source terms. Hence, from now on, we assume $q$ to be of the form $q = i \hat{q}, \, \hat{q} \in \mathbb{R}_{>0}$.

If we first apply the transformation $q \rightarrow i \hat{q}$ and than use equalities (\ref{RES_I}) and (\ref{RES_II}), the integral (\ref{INT_I}) becomes
\begin{equation} 
 \frac{1}{2 \pi i} \oint_{\gamma}
 \, e^{- D \hat{q}^2 t} \, \frac{g(x,i \hat{\hat{q}}|x_0)}{h(i \hat{q})} \, 2 D i \hat{q} \, d\hat{q} = \sum_{i=1}^{\infty} e^{ - D \hat{q}^2 t} \, \left. \frac{g(x,i \hat{q}|x_0)}{\partial_q h(i \hat{q})} \right|_{\hat{q} = \beta_i} \, 2 D i \hat{q}
\end{equation}
Now we can fill in our Laplace solutions and obtain the final solution in real space
\begin{multline}
 p_r(x,t|x_0) = -2 \sum_{i=1}^{\infty} \, e^{- D \beta_i ^2 t} \, \mathrm{sin} \, \beta_i (L_r - x) \\ 
\frac{D \beta_i \, \mathrm{sin} \, \beta_i (L_l + x_0) + k \, \mathrm{sin} \, \beta_i L_l \,\, \mathrm{sin} \, \beta_i x_0}{D \, (L \beta_i \mathrm{cos} \, \beta_i L + \, \mathrm{sin} \, L \beta_i ) + k \,( L_r \, \mathrm{cos} \, \beta_i L_r \,\, \mathrm{sin} \, \beta_i L_l + L_l \, \mathrm{cos} \, \beta_i L_l \,\, \mathrm{sin} \, \beta_i L_r )},
 \label{FSOL_R}
\end{multline}

\begin{multline}
 p_l(x,t,|x_0) = -\sum_{i=1}^{\infty} e^{- D \beta_i ^2 t} \, 2 D \beta_i \, \\ 
\frac{\mathrm{sin} \, \beta_i (L_l + x) \,\, \mathrm{sin} \, \beta_i (L_r - x_0)}{D \, (L \beta_i \mathrm{cos} \, \beta_i L + \, \mathrm{sin} \, L \beta_i ) + k \,( L_r \, \mathrm{cos} \, \beta_i L_r \,\, \mathrm{sin} \, \beta_i L_l + L_l \, \mathrm{cos} \, \beta_i L_l \,\, \mathrm{sin} \, \beta_i L_r )},
\label{FSOL_L}
\end{multline}
where $L=L_r + L_l$ in both equations. Note that when $p_r$ is evaluated for $x<x_0$, $x$ and $x_0$ have to be interchanged in equation (\ref{FSOL_R}). This is due to the absolute sign in the Ansatz (\ref{Anz_R}).

\subsubsection{Solution for a sink in an unbounded domain}

For completes and its usefulnesses in $\GFRD$, we end this derivation with the Green's function for an absorbing sink at $x_s$ in the absence of absorbing boundaries at both sides
\begin{multline}
 p(x,t|x_0) = \frac{1}{\sqrt{4 \pi D t}}e^{\frac{-(x-x_0)^2}{4 D t}} - \frac{k}{4 D}e^{\frac{k}{2 D}(|x-x_s|+|x_s-x_0|)+\frac{k^2 t}{4 D}} \\
\mathrm{erfc}\left( \frac{|x-x_s|+|x_s-x_0|}{\sqrt{4 D t}}+\frac{k}{2D}\sqrt{D t} \right).
\label{FSOL_NB}
\end{multline}
We note that we can map this solution to the solution of a radiation boundary condition at $x_s$ in a semi-infinite domain when we change $k \rightarrow 2 k$ and multiply the second term with 1/2. The servival probility function $S(t|x0)$ of the above function is equal to that of the radiation boundary when we double the rate $k$.

\subsection{Important quantities for $\GFRD$}

For the functionallity of $\GFRD$ we need the survival probability $S(t|x_0)$, the cumulative distribution function (c.d.f) in space $F(x^*,t|x_0)$ and the fluxes $q_x(t|x_0)$ leaving the domain. Probability flux leaves the domain at the left and right boundaries and via the sink.

Remeber that the Green's functions (\ref{SOL_R}) and (\ref{SOL_L}) were of the form
\begin{equation} 
 \sum_{i=1}^{\infty} e^{-D \beta_i ^2 t} \, \frac{g_{r/l}(x, \beta_i |x_0)}{ h'(\beta_i)}
\end{equation}
where $h'$ denotes the derivative of the denominator function $h(q)$. Because the required quantaties are obtained by integrating or differentiating the Green's functions with respect to their spatial coordinate $x$, the only part of the solution which changes is the numerator function $g_{r/l}$. Therefore, below we only write this part the equation.

The survival probability is defined by the spatial integral of the Green's function over the whole domain. Because there is only a source (the initial delta peak) at $t = 0$, we expect $S(t)$ to be a monotonic decaying fucntion starting at 1,
\begin{equation} 
 S(t|x_0) = \int_{-L_l}^{0} p_l(x,t|x_0) \, dx \, + \, \int_{0}^{L_r} p_r(x,t|x_0) \, dx.
\end{equation}
The c.d.f. for the spatial coordinate $x$ is the integral of the Green's function from $-L_l$ to an arbitrary position $x^* \in [-L_l,L_r]$. Because we want the c.d.f. to be normelized, a.k. the c.d.f. $F \rightarrow 1$ as $x^* \rightarrow L_r$, we devide the integral with the survival probability,
\begin{equation}
 \begin{split}
  F(x^*,t|x_0) & = \frac{1}{S(t|x_0)} \int_{-L_l}^{x*} p_l(x,t|x_0) \, dx \quad (x^* \leq 0), \\
  F(x^*,t|x_0) & = \frac{1}{S(t|x_0)} \left( \int_{-L_l}^{0} p_l(x,t|x_0) \, dx \, + \, \int_{0}^{x^*} p_r(x,t|x_0) \, dx. \right) \quad (x^* > 0) 
 \end{split}
\end{equation}
The flux (net current moving to the right) through a point $x'$ lying within the domain is,

\begin{equation}
 q(t|x_0) = -D \left. \partial_x p(x,t|x_0) \right|_{x = x'}
\end{equation}

\subsection{Numerical evaluation}

In order for the root fuction to be easily evaluated by the rootfinder, we apply the transformations $L_l + L_r \rightarrow L$ and $L_r - L_l \rightarrow L'$ and let $x = s L$, and obtain,
\begin{equation}
 x \, \mathrm{sin} \, x \, + \, \frac{k \, L}{2 \, D} \, \left( \, \mathrm{cos} \, x \frac{L'}{L} \, - \, \mathrm{cos} \, x \right) = 0
\end{equation}
for the denominator of the Green's functions $\partial_q h(q)$, after $q \rightarrow i s$ we write with the same transformations as above,
\begin{equation}
 D \left( L s \, \mathrm{cos} \, s L \, + \, \mathrm{sin} \, s L \, \right) \, + \, \frac{k}{2} \, \left( L \, \mathrm{sin} \, s L \, - \, L' \, \mathrm{sin} \, s L' \right).
\end{equation}
The numerator for the survival probability $S(t|x_0)$ is found to equal
\begin{multline}
2 D \left[ \mathrm{sin} \, q L \, - \, \mathrm{sin} \, q (L_r - x_0) \, - \, \mathrm{sin} \, q (L_l + x_0)  \right] \, + \, \\ 
 \frac{2 k}{q} \mathrm{sin} \, q L_l \left[ \mathrm{sin} \, q L_r \, - \, \mathrm{sin} \, q x_0 \, - \, \mathrm{sin} \, q (L_r - x_0) \right] 
\end{multline}